A Technique for Inter-Retailer Recommendations

### 2014-06-12 paper2/paper.tex
Minor corrections.

\section*{Abstract}
\emph{The problem of making recommendation across retailers using only publicly
available information is explored. Several inter-retailer recommender algortihms
are presented. Each is compared in an experiment conducted with real data. A 
novel latent feature recommender that leverages public intra-retailer
recommendation information is described and shown to outperform a nontrivial
content-based approach.}

@@

\subsubsection*{Extracting Latent Features}
We may now construct a synthetic $N \times N$ item-user matrix $M$ with rows
representing items and columns representing users, which are in fact the
stationary distributions discovered from our random walks. This matrix has at
least one notable advantage over most real item-user matrices---it is not
sparse.  Rather, each synthetic user has a finite value associated with every
item that is reachable from its origin item in the graph. Unlike the typical CF
scenario, we may therefore employ SVD as our matrix factorization technique
since our item-user matrix is virtually complete (i.e. nearly void of empty
entries).  Using SVD, we factor our item-user matrix as
\begin{align}
M = U \Sigma V^T,
\end{align}
where the rows of the $N \times N$ matrix $U$ and the $N \times N$ matrix $V$
are the items and users represented in an $N$-dimensional concept space,
respectively, and the $N$ diagonal entries of $\Sigma$ are the singular values
associated with each concept.

@@

Figures 1 and 2 show the top and bottom most relevant items for two of the sixteen
concepts learned in this way on the Women's Dresses category. Note that concepts
learned by SVD range from positive to negative, representing the two extremes of
the concept, therefore it is relevant to look at both the top and bottom items.

@@

\subsubsection*{Concept Space Mapping}
Using random walks to simulate user data, followed by the application of LSI, is
indeed an effective way to encode information from a recommendation graph into a
latent feature space. But what is the relationship between latent feature
vectors derived from disconnected graph partitions? In principle, there is no
guarantee of any relationship between such concept spaces in the same way that
there is no guarantee of any relationship between two arbitrary, disconnected
recommendation graphs. Nonetheless, our hypothesis is that a relationship
between these concept spaces will often exist due to the effectiveness of the
underlying CF techniques that generate the graphs to capture meaningful
item-item associations. Put another way, we believe that the most fundamental
latent dimensions along which products in a category such as Women's swimsuits
vary will be common to many swimsuit catalogues. If this is indeed true, then it
might be possible to learn a mapping between these concept spaces.

If a mapping between the concept spaces associated with two catalogues were
known, we could translate all items into a common space and then choose
inter-catalogue edges between item pairs that are nearest to one another in
concept space. To learn these mappings, we must revisit the notion of content
similarity. Once again, we utilize tf-idf vectors. This time, however, we weight
an item $i$'s tf-idf vector $\vec{\tau}(i)$ by the elements in its latent
feature vector $\vec{v}(i)$. In this way, we construct a dense {\em
super-tf-idf} vector for each concept of each catalogue. Formally, we say that
a concept $j$ of a catalogue $C$ is represented by the super-tf-idf vector
\begin{align}
\vec{s}(j,C) = \sum_{i \in C}{v(i)_j \cdot \vec{\tau}(i)},
\end{align}
where $v(i)_j$ is the $j$\textsuperscript{th} component of item $i$'s latent
feature vector. Note that since the components $v(i)_j$ may be positive or
negative so too can the elements of a concept's super-tf-idf vector, in
constrast to the elements of an individual item's tf-idf vector which are
necessarily greater than or equal to zero. Also note that we compute both
$\vec{s}_1$ and $\vec{s}_2$, which correpond to unigram and bigram super-tf-idf
vectors, respectively, for each concept. These super-tf-idf vectors then act as
proxies for the concept.

With these proxy representations of the concepts for the concept spaces of two
catalogues $C_1$ and $C_2$, we construct a matrix $A$ of inner-products of
super-tf-idf vectors.  Specifically, the matrix elements of $A$ are given by
\begin{align}
A_{ij}(C_1,C_2) = \vec{s}_1(i,C_1) \cdot \vec{s}_1(j,C_2) + \beta \cdot
\vec{s}_2(i,C_1) \cdot \vec{s}_2(j,C_2),
\end{align}
where $i$ is a concept from $C_1$ and $j$ is a concept from $C_2$ and $\beta$ is
the same parameters as before (again, we find $\beta=0.3$ works well).  We may
now use this matrix to transform the vector representation of an item in one
concept space into a vector representation in another concept space.

It is important to note that the transformation matrix $A$ is not orthogonal, and
therefore, does not preserve vector norms. We address this issue by normalizing
all feature vectors, both foreign and domestic, in the concept space followingj
the transformation. This radially projects all items onto the surface of a
hypersphere in the 16-dimensional concept space. Now, when ranking an item's
neighbors by distance, we are actually ranking neighbors by angular separation.
That is, we are ranking neighbors by their cosine similarity in concept space.
After experimenting with various normalization schemes we find that this
approach works best.

@@

Next, we reincorporate popularity by modifying how we search for nearest
neighbors in concept space. Instead of simply ranking neighbors by distance, we
now rank by distance divided by $\rho$.  In this context, having defined
popularity as the logorithm of node degree reduces the liklihood that one or two
very popular items will always appear as the best recommendations when using
this weighted ranking policy.

The third way in which popularity is used is determining the number
recommendations alotted per product. We assume that a popular product in one
catalogue is likely to be a popular product in another catalogue. Specifically,
the number of predicted inter-catalogue recommendations predicted per item is
made directly proportional to the item's value of $\rho$.

>>>>

\section*{Abstract}
\emph{The problem of making recommendation across retailers using only publicly
available information is explored. Several inter-retailer recommender algorithms
are presented. Each is compared in an experiment conducted with real data. A 
novel latent feature recommender that leverages public intra-retailer
recommendation information is described and shown to outperform a nontrivial
content-based approach.}

@@

\subsubsection*{Extracting Latent Features}
We may now construct a synthetic $N \times N$ user-item matrix $M$ with rows
representing items and columns representing users, which are in fact the
stationary distributions discovered from our random walks. This matrix has at
least one notable advantage over most real user-item matrices---it is not
sparse.  Rather, each synthetic user has a finite value associated with every
item that is reachable from its origin item in the graph. Unlike the typical CF
scenario, we may therefore employ SVD as our matrix factorization technique
since our user-item matrix is virtually complete (i.e. nearly void of empty
entries).  Using SVD, we factor our user-item matrix as
\begin{align}
M = U \Sigma V^T,
\end{align}
where the rows of the $N \times N$ matrix $U$ and the $N \times N$ matrix $V$
are the items and users represented in an $N$-dimensional concept space,
respectively, and the $N$ diagonal entries of $\Sigma$ are the singular values
associated with each concept.

@@

Figures 1 and 2 show the top and bottom most relevant items for two of the sixteen
concepts learned for the Women's Dresses category. Note that concepts learned by
SVD range from positive to negative, representing the two extremes of the
concept.

@@

\subsubsection*{Concept Space Mapping}
Using random walks to simulate user data, followed by the application of LSI, is
indeed an effective way to encode information from a recommendation graph into a
latent feature space. But what is the relationship between latent feature
vectors derived from disconnected graph partitions? In principle, there is no
guarantee of any relationship between such concept spaces in the same way that
there is no guarantee of any relationship between two arbitrary, disconnected
recommendation graphs. Nonetheless, our hypothesis is that a relationship
between these concept spaces often will exist due to the effectiveness of the
underlying CF techniques that generate the graphs to capture meaningful
item-item associations. Put another way, we believe that the most fundamental
latent dimensions along which products in a category such as Women's swimsuits
vary will be common to many swimsuit catalogues. If this is indeed true, then it
might be possible to learn a mapping between these concept spaces.

If a mapping between the concept spaces associated with two catalogues were
known, we could translate all items into a common space and then choose
inter-catalogue edges between item pairs that are nearest to one another in
concept space. To learn these mapping, we must revisit the notion of content
similarity. Once again, we utilize tf-idf vectors. This time, however, we weight
an item $i$'s tf-idf vector $\vec{\tau}(i)$ by the elements in its latent
feature vector $\vec{v}(i)$. In this way, we construct a dense {\em
super-tf-idf} vector for each concept for each catalogue. Formally, we say that
a concept $j$ of a catalogue $C$ is represented by the super-tf-idf vector
\begin{align}
\vec{s}(j,C) = \sum_{i \in C}{v(i)_j \cdot \vec{\tau}(i)},
\end{align}
where $v(i)_j$ is the $j$\textsuperscript{th} component of item $i$'s latent
feature vector. Note that since the components $v(i)_j$ may be positive or
negative so too can the elements of a concept's super-tf-idf vector, in
constrast to the elements of an individual item's tf-idf vector which are
necessarily greater than or equal to zero. Also note that we compute both
$\vec{s}_1$ and $\vec{s}_2$, which correpond to unigram and bigram super-tf-idf
vectors, respectively, for each concept. These super-tf-idf vectors then act as
proxies for the concept.

With these proxy representations of the concepts for the concept spaces of two
catalogues $C_1$ and $C_2$, we construct a matrix $A$ of inner-products of
super-tf-idf vectors.  Specifically, the matrix elements of $A$ are given by
\begin{align}
A_{ij}(C_1,C_2) = \vec{s}_1(i,C_1) \cdot \vec{s}_1(j,C_2) + \beta \cdot
\vec{s}_2(i,C_1) \cdot \vec{s}_2(j,C_2),
\end{align}
where $i$ is a concept from $C_1$ and $j$ is a concept from $C_2$ and $\beta$ is
the same parameter as before (again, we find $\beta=0.3$ works well).  We may
now use this matrix to transform the vector representation of an item in one
concept space into a vector representation in another concept space.

It is important to note that the transformation matrix $A$ is not orthogonal, and
therefore, does not preserve vector norms. We address this issue by normalizing
all feature vectors, both foreign and domestic, in the concept space following
the transformation. This radially projects all items onto the surface of sphere
within the 16-dimensional concept space. Now when ranking an item's neighbors by
distance, we are actually ranking neighbors by angular separation. That is, we
are ranking neighbors by their cosine similarity in concept space. After
experimenting with various normalization schemes we find that this approach
works best.

@@

Next, we reincorporate popularity by modifying how we search for nearest
neighbors in concept space. Instead of simply ranking neighbors by distance, we
now rank by distance divided by $\rho$.  In this context, having defined
popularity as the logorithm of node degree, instead of simply node degree,
reduces the liklihood that one or two very popular items will always appear as
the best recommendations when using this weighted ranking policy.

The third way in which popularity is used is determining the number of
recommendations alotted per product. We assume that a popular product in one
catalogue is likely to be a popular product in another catalogue. Specifically,
the number of predicted inter-catalogue recommendations predicted per item is
made directly proportional to the item's value of $\rho$.
